Reproducible and Programmatic Generation of Neuroimaging Visualizations

### 2022-08-15 manuscript.Rmd
c++ compline edit
python fig draft added

---
title: Reproducible and Programmatic Generation of Neuroimaging Visualizations
authors:
  - name: Sidhant Chopra
    thanks: Corresponding author
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
    email: sidhant.chopra@yale.com
  - name: Loïc Labache
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Elvisha Dhamala
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Edwina R Orchard
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Avram Holmes
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
keywords:
  - Neuroimaging visualization
  - Reproducibility 
  - Programmatic figures
  - Open Science
  - R
  - Python
  - Brain Visualization
abstract: Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results. Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI). There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB. Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process. This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments. Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data.
output: 
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
    keep_tex: true
    latex_engine: lualatex
  bookdown::word_document2:
    toc: false
linestretch: 2
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{setspace}\doublespacing
---

@@

Increasingly popular software such as *R Markdown*, *Quarto* and *Jupyter Notebook* allow for the mixing of prose and code in a single script, resulting in fully reproducible and publication ready papers. By using code-based tools available within your preferred environment, brain visualizations can be directly integrated and embedded within a paper or report. For instance, a fully reproducible version of the current paper can be found on [GitHub](https://github.com/sidchop/Brain_comms_field_potential). Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g. *e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code computational environment linked to the associated article using 'compute capsules' from the cloud-based platform Code Ocean (Code Ocean, 2021).

@@

|:-----------------------|:--------|:--------|:--------------|:--------|:------------|

@@

| ANTsR                  | ***+*** | ***+*** | ***+***       |         |             |
| brainconn              |         |         |               | ***+*** |             |
| brainR                 | ***+*** |         | ***+***       |         |             |
| ciftitools             | ***+*** | ***+*** | ***+*****\*** |         |             |
| fsbrain                | ***+*** | ***+*** | ***+*****\*** |         |             |

@@

# Examples of Neuroimaging Visualization Packages available in *R*

The following four sections provide brief examples of packages and functions available in R for visualizing voxel, vertex, ROI and edge-level data. A didactic introduction to these tools is provided in an accompanying [collection of beginner friendly R-Notebooks](https://github.com/sidchop/neuRo-vis-BrainHack-2021). This is not an exhaustive list of packages available for visualizing brain data in R, rather, the following sections and accompanying notebooks aim to give the reader a sense of the available options, and an entry point to using them.

@@

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations in R') &
    theme(plot.tag = element_text(face = 'bold'),
          plot.title = element_text(face="bold"))

@@

*Figure 1. Examples of brain imaging visualization make in using different R-packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown. Made using the the `ortho2` function from the `neurobase` package. B) A voxel-level cortical parcellation overlaid on T1-weighted MRI data shown in 9-slice axial orientation. Made using the `overlay` function from the `neurobase` package. C) A CIFTI format surface ROI atlas with a corresponding statistic assigned to each regions, with both hemispheres displayed on a inflated template surface in lateral view. Made using the `view_xifti_surface` from the `ciftiTools` package. D) A coronal cross-sectional rendering of subcortical structures where a statistical value has been assigned to each region. Made using the `aseg` atlas from the `ggseg` package. E) A 3D rendering of 9 bilateral subcortical regions where a statistical value has been assigned to each region. Made using the `aseg` atlas from the `ggseg3d` package. F) Medial and lateral views of a ROI atlas displayed on inflated template cortical surface where a statistical value has been assigned to each region. Made using the `glasser` atlas from the `ggsegGlasser` package, which was plotted using `ggseg`. G) Lateral view of a CIFTI format vertex-level data displayed on a inflated template surface. Made using the `view_xifti_surface` function from the `ciftiTools` package. H) Medial and lateral views of vertex-level data displayed on a individuals white matter surface. Made using the `vis.subject.morph.standard` function from the `fsbrain` package. I) A weighted and undirected graph plotted on top, left and front views of a template of a brain rendering. Made using the `brainconn` function from the `brainconn` package. All code used to compile this figure, as well as the contents of each panel are provided in the Supplement.

>>>>

---
title: Reproducible and Programmatic Generation of Neuroimaging Visualizations
authors:
  - name: Sidhant Chopra
    thanks: Corresponding author
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
    email: sidhant.chopra@yale.com
  - name: Loïc Labache
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Elvisha Dhamala
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Edwina R Orchard
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Avram Holmes
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
keywords:
  - Neuroimaging visualization
  - Reproducibility 
  - Programmatic figures
  - Open Science
  - R
  - Python
  - Brain Visualization
abstract: Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results. Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI). There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB. Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process. This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments. Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data.

output: 
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
    keep_tex: true
    latex_engine: lualatex
  bookdown::word_document2:
    toc: false
linestretch: 2
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{setspace}\doublespacing
---

@@

Increasingly popular software such as *R Markdown*, *Quarto* and *Jupyter Notebook* allow for the mixing of prose and code in a single script, resulting in fully reproducible and publication ready papers. By using code-based tools available within your preferred environment, brain visualizations can be directly integrated and embedded within a paper or report. For instance, a fully reproducible version of the current paper can be found on [GitHub](https://github.com/sidchop/RepoNeuroVis). Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g. *e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code computational environment linked to the associated article using 'compute capsules' from the cloud-based platform Code Ocean (Code Ocean, 2021).

@@

|:-----------|:-----------|:-----------|:-----------|:-----------|:-----------|

@@

| [ANTsR](https://github.com/ANTsX/ANTsR)                | ***+*** | ***+*** | ***+***       |         |             |
| [brainconn](https://github.com/sidchop/brainconn)      |         |         |               | ***+*** |             |
| [brainR](https://github.com/muschellij2/brainR)        | ***+*** |         | ***+***       |         |             |
| [ciftitools](https://github.com/mandymejia/ciftiTools) | ***+*** | ***+*** | ***+*****\*** |         |             |
| [fsbrain](https://github.com/dfsp-spirit/fsbrain)      | ***+*** | ***+*** | ***+*****\*** |         |             |

@@

# Examples of Neuroimaging Visualization Packages available in *R and Python*

The following four sections provide brief examples of packages and functions available in R and Python for visualizing voxel, vertex, ROI and edge-level data. This is not an exhaustive list of packages available for visualizing brain data in R and Python, rather, the following sections aim to give the reader a sense of the available options, and an entry point to using them. A didactic introduction to the R based tools is provided in an accompanying [collection of beginner friendly R-Notebooks](https://github.com/sidchop/neuRo-vis-BrainHack-2021).

@@

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations made in R') &
    theme(plot.tag = element_text(face = 'bold'),
          plot.title = element_text(face="bold"))

@@

*Figure 1. Examples of brain imaging visualization make in using different R packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown. Made using the the `ortho2` function from the `neurobase` package. B) A voxel-level cortical parcellation overlaid on T1-weighted MRI data shown in 9-slice axial orientation. Made using the `overlay` function from the `neurobase` package. C) A CIFTI format surface ROI atlas with a corresponding statistic assigned to each regions, with both hemispheres displayed on a inflated template surface in lateral view. Made using the `view_xifti_surface` from the `ciftiTools` package. D) A coronal cross-sectional rendering of subcortical structures where a statistical value has been assigned to each region. Made using the `aseg` atlas from the `ggseg` package. E) A 3D rendering of 9 bilateral subcortical regions where a statistical value has been assigned to each region. Made using the `aseg` atlas from the `ggseg3d` package. F) Medial and lateral views of a ROI atlas displayed on inflated template cortical surface where a statistical value has been assigned to each region. Made using the `glasser` atlas from the `ggsegGlasser` package, which was plotted using `ggseg`. G) Lateral view of a CIFTI format vertex-level data displayed on a inflated template surface. Made using the `view_xifti_surface` function from the `ciftiTools` package. H) Medial and lateral views of vertex-level data displayed on a individuals white matter surface. Made using the `vis.subject.morph.standard` function from the `fsbrain` package. I) A weighted and undirected graph plotted on top, left and front views of a template of a brain rendering. Made using the `brainconn` function from the `brainconn` package. All code used to compile this figure, as well as the contents of each panel are provided in the Supplement.

```{=tex}
\newpage
\blandscape
```
```{r Figure2 , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center'}

fig_name <- c('fig1', 'fig2', 'fig3', 'fig4', 'fig5', 'fig6', 'fig7')
figs <- list()

for (f in 1:length(fig_name)) { 
  knitr::plot_crop(paste0('data/python/',fig_name[f],'.png'))
  figs[[f]] <- ggplot() +
    annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/python/',fig_name[f],'.png')),
                                                             width=ggplot2::unit(1,"npc"),
                                                             height=ggplot2::unit(1,"npc")))
}

patchwork2 <- (figs[[1]]+figs[[3]])/figs[[2]] | (figs[[6]]/figs[[4]]/figs[[5]])

  figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]])+plot_layout(heights = c(0.5,1,1)))

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations made in Pythoin') &
    theme(plot.tag = element_text(face = 'bold'),
          plot.title = element_text(face="bold"))

*Figure 1. Examples of brain imaging visualization make in using different Python packages.* A)
